---
title: "scRNAseq"
author: "Ayshwarya Subramanian"
date: "9/15/2017"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE)

```
## scRNAseq datasets

For the application of differential gene expression, we look at scRNAseq and bulk RNAseq data. Here we discuss the analyses of scRNAseq.

```{r echo=FALSE, results='asis'}
library(knitr)
df=data.frame(datasets=c("GSE63818-GPL16791","GSE45719"),organism=c("human","mouse"),technology=c("smart-seq","smart-seq"))
kable(df,caption="scRNAseq Datasets")
df1=data.frame(DE_method=c("Wilcoxon-rank sum","ScDD","MAST"),input=c("tpm","count","tpm"),output=c("p-val","p-val","p-val"))
kable(df1,caption="DE methods")

```

We selected two datasets, one each from mouse and human spp, from the `Conquer` (Soneson et al., 2017) scRNAseq data collection. `Conquer` provides uniformly processed gene expression data summarized to both count (aggregated from transcript members), and length-scaled TPMs at the gene level. The authors use `salmon` to quantify raw reads and `tximport` to summarize to gene level. The data comes from Smartseq, a plate based sequencing method which results in deeper sequencing of reads at the cost of numbers of cells. 

We also select two droplet-based datasets from the `10x` website:4k PBMC and 4k T cells, both from healthy donors.  

## Data Download

We first download the data in our working directory. First, let's download the Human data:GSE63818-GPL16791 

```{bash, humandata-download, eval=FALSE}
# all data and results are on the wd: /net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/fdrbenchmark-scrnaseq

#these look well separated on the tsne plots on the conquer manuscript supplement
wget http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE63818-GPL16791.rds
# qc report: 
wget http://imlspenticton.uzh.ch/robinson_lab/conquer/report-multiqc/GSE63818-GPL16791_multiqc_report.html
# scater report
wget http://imlspenticton.uzh.ch/robinson_lab/conquer/report-scater/GSE63818-GPL16791_scater.html
```

Next, we download the Mouse data:GSE45719

```{bash, mousedata-download, eval=FALSE}
#another well-separated dataset on tsne: GSE60749−GPL13112
wget http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE45719.rds
```

We directly download the cell-ranger processed filtered expression matrices [4k PBMC cells](https://support.10xgenomics.com/single-cell-gene-expression/datasets/2.1.0/pbmc4k) and [4k T-cells](https://support.10xgenomics.com/single-cell-gene-expression/datasets/2.1.0/t_4k) from the 10x website.

## R workspace setup

The downloaded plate data are stored in R objects called `MultiAssayExperiment`; this package requires BioConductor release 3.5 and R version 3.4 or greater for installation. 

```{r, workspace-setup, eval=FALSE}
#installation
# source("https://bioconductor.org/biocLite.R")
# biocLite("swfdr")
# biocLite("IHW")
# biocLite("qvalue")
# install.packages("ashr")
# install.packages("fdrtool")
# biocLite("edgeR")
# source("https://bioconductor.org/biocLite.R")
# biocLite("scran")
#  Samseq installation
# source("https://bioconductor.org/biocLite.R")
# biocLite("impute") 
# install.packages("samr")
# source("https://bioconductor.org/biocLite.R")
#biocLite("MultiAssayExperiment")
#source("https://bioconductor.org/biocLite.R")
#biocLite("SummarizedExperiment")
#biocLite("BiocGenerics")
#biocLite("S4Vectors")
#biocLite("BiocParallel")
#install_github("areyesq89/SummarizedBenchmark")
#biocLite("EBSeq")
#biocLite("scDD")
#source("https://bioconductor.org/biocLite.R")
#biocLite("SingleCellExperiment")
#source("https://bioconductor.org/biocLite.R")
#biocLite("MAST")
#install.packages("FDRreg")

# packages for data analyses
library(MultiAssayExperiment)
library(dplyr)
library(ggplot2)
library(cowplot)
library(UpSetR)
library(hexbin)

# packages for differential expression testing
library(impute)
#library(samr)
library(scDD)
library(EBSeq)
library(edgeR)
library(MAST)

# benchmark methods
library(IHW)
library(ashr)
library(qvalue)
library(swfdr)
library(fdrtool)
library(FDRreg)

# summarizebenchmark
library(SummarizedBenchmark)
library(SingleCellExperiment)
source("/Users/subraman/Documents/research/fdr_jc/code/fdr-benchmark-scrnaseq-helper.R")

```

## Data processing and filtering

For data processing, we follow practices and recommendations from Soneson et al., 2017. Briefly, we limit ourselves to 2-group comparisons, and  select the maximal size defined as the smallest number of cells in either group as the group-level sample size for the DE analyses. 

First, we get the human data for the two groups of interest.

We look at results with and without prefiltering genes (keeping genes with TPM > 1 in atleast 25% of cells)

```{r, dataload, eval=FALSE}
setwd("/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/fdrbenchmark-scrnaseq")
setwd("/Users/subraman/Documents/research/fdr_jc/data/")
#load compressed mae
humandata=readRDS("GSE63818-GPL16791.rds")
#get gene-level data
genes=experiments(humandata)[1]
#TPMS
#get tpms
tpms=assays(genes$gene)[1]
#get tpm dataframe
human_tpms=tpms$TPM
# get column data/metadata for sample
cdata=colData(humandata)
cdata=as.data.frame(cdata)
# select groups of interest
g1 = cdata%>%subset(source_name_ch1=="Somatic Cells" & characteristics_ch1=="developmental stage: 7 week gestation")
g2 = cdata%>%subset(source_name_ch1=="Primordial Germ Cells" & characteristics_ch1=="developmental stage: 7 week gestation")
# select maximal sample size
sample_size = c(dim(g1)[1],dim(g2)[1])
ssize = which.min(sample_size)
g1_ids=sample.int(dim(g1)[1], size = sample_size[ssize], replace = FALSE)
g2_ids=sample.int(dim(g2)[1], size = sample_size[ssize], replace = FALSE)
# get group data of maximal size
g1_data=human_tpms[,colnames(human_tpms)%in%g1$geo_accession[g1_ids]]
g2_data=human_tpms[,colnames(human_tpms)%in%g2$geo_accession[g2_ids]]
#save data
saveRDS(g1_data,"h_g1_data.rds")
saveRDS(g2_data,"h_g2_data.rds")
#COUNTS
#get counts
count=assays(genes$gene)[2]
#get count dataframe
human_count=count$count
# get column data/metadata for sample
cdata=colData(humandata)
cdata=as.data.frame(cdata)
# get group data of maximal size
g1_cdata=human_count[,colnames(human_count)%in%g1$geo_accession[g1_ids]]
g2_cdata=human_count[,colnames(human_count)%in%g2$geo_accession[g2_ids]]
#save data
saveRDS(g1_cdata,"h_g1_cdata.rds")
saveRDS(g2_cdata,"h_g2_cdata.rds")

```

For the human data, each group has 26 cells and 65218 genes. We merge the data and remove all genes which have 0s in all samples

```{r, nofilter, eval=FALSE}
g_data=cbind(g1_data,g2_data)
g_data_sub=g_data[rowSums(g_data==0)!=52,] # 0.25*52=13
saveRDS(g_data_sub,"h_g_nofilter_subdata.rds") # removing rows with all zeros reduces the number of genes to 33169

g_cdata=cbind(g1_cdata,g2_cdata)
g_cdata_sub=g_cdata[rowSums(g_cdata==0)!=52,]
saveRDS(g_cdata_sub,"h_g_nofilter_subcdata.rds") # removing rows with all zeros reduces the number of genes to 33169

saveRDS(g1_ids,"g1_ids.rds")
saveRDS(g2_ids,"g2_ids.rds")
```

```{r,mouse}
##mouse data
mousedata=readRDS("GSE45719.rds")
#get gene-level data
genes=experiments(mousedata)[1]
#TPMS
#get tpms
tpms=assays(genes$gene)[1]
#get tpm dataframe
mouse_tpms=tpms$TPM
# get column data/metadata for sample
cdata=colData(mousedata)
cdata=as.data.frame(cdata)
# select groups of interest
g1 = cdata%>%subset(source_name_ch1=="16-cell stage blastomere")
g2 = cdata%>%subset(source_name_ch1=="Mid blastocyst cell (92-94h post-fertilization)")
# select maximal sample size
sample_size = c(dim(g1)[1],dim(g2)[1])
ssize = which.min(sample_size)
g1_ids=sample.int(dim(g1)[1], size = sample_size[ssize], replace = FALSE)
g2_ids=sample.int(dim(g2)[1], size = sample_size[ssize], replace = FALSE)
# get group data of maximal size
g1_data=mouse_tpms[,colnames(mouse_tpms)%in%g1$geo_accession[g1_ids]]
g2_data=mouse_tpms[,colnames(mouse_tpms)%in%g2$geo_accession[g2_ids]]
#save data
saveRDS(g1_data,"m_g1_data.rds")
saveRDS(g2_data,"m_g2_data.rds")
#COUNTS
#get counts
count=assays(genes$gene)[2]
#get count dataframe
mouse_count=count$count
# get column data/metadata for sample
cdata=colData(mousedata)
cdata=as.data.frame(cdata)
# get group data of maximal size
g1_cdata=mouse_count[,colnames(mouse_count)%in%g1$geo_accession[g1_ids]]
g2_cdata=mouse_count[,colnames(mouse_count)%in%g2$geo_accession[g2_ids]]
#save data
saveRDS(g1_cdata,"m_g1_cdata.rds")
saveRDS(g2_cdata,"m_g2_cdata.rds")

```

For the mouse data, each group has 50 cells and 45686 genes. We merge the data and remove all genes which have 0s in all samples

```{r, nofilter, eval=FALSE}
g_data=cbind(g1_data,g2_data)
g_data_sub=g_data[rowSums(g_data==0)!=dim(g_data)[2],] # 0.25*52=13
saveRDS(g_data_sub,"m_g_nofilter_subdata.rds") # removing rows with all zeros reduces the number of genes to 33169

g_cdata=cbind(g1_cdata,g2_cdata)
g_cdata_sub=g_cdata[rowSums(g_cdata==0)!=dim(g_data)[2],]
saveRDS(g_cdata_sub,"m_g_nofilter_subcdata.rds") # removing rows with all zeros reduces the number of genes to 33169

saveRDS(g1_ids,"m_g1_ids.rds")
saveRDS(g2_ids,"m_g2_ids.rds")
```


## Differential expression testing

We perform differential expression testing using 3 methods on the unfiltered data:

### Wilcoxon rank-sum test

Per `Conquer`, we run TMM normalization to the TPM units before running Wilcoxon.
```{r, wilcoxon, eval=FALSE}
g_data_sub=readRDS("h_g_nofilter_subdata.rds")
groups=c(rep(1,26),rep(2,26))
compute_wilcoxde_sc(g_data_sub,"h_data_wilcox_p.rds",groups)

g_data_sub=readRDS("m_g_nofilter_subdata.rds")
groups=c(rep(1,50),rep(2,50))
compute_wilcoxde_sc(g_data_sub,"m_data_wilcox_p.rds",groups)
## warnings:
## In wilcox.test.default(x = structure(c(0, 0, 14.2107316559397,  ... :
## cannot compute exact p-value with ties

```

### Differences in distributions of gene expression in the 2 conditions

We apply scDD to log-transformed counts as below:

```{r, scdd, eval=FALSE}
# library(SummarizedExperiment)
# condition=c(rep(1,26),rep(2,26))
# rownames(ng_cdata_sub) <- paste0(rownames(ng_cdata_sub), 1:nrow(ng_cdata_sub), sep="")
# colnames(ng_cdata_sub) <- names(condition) <- paste0("Sample",
# 1:ncol(ng_cdata_sub), sep="")
# SDSumExp <- SummarizedExperiment(assays=list("NormCounts"=ng_cdata_sub),
# colData=data.frame(condition))
#saveRDS(SDSumExp,"data_sumexp.rds")

## human data
compute_scdd_sc(g_cdata_sub,"h_sizefactors.rds","h_data_sce.rds","humanplate_scDatEx_ks.rds")
##Notice: 16096 genes have less than 3 nonzero cells per condition.  Skipping these genes.
##Clustering observed expression data for each gene
##Setting up parallel back-end using 6 cores
# perm=1000
# scDatExperm <- scDD(SDSumExp, prior_param=prior_param, param=param, permutations=1000,testZeroes=FALSE)
# saveRDS(scDatExperm,"/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/fdrbenchmark-scrnaseq/scDatExperm.rds")

## mouse data
compute_scdd_sc(g_cdata_sub,"m_sizefactors.rds","m_data_sce.rds","mouseplate_scDatEx_ks.rds")
# Notice: 10246 genes have less than 3 nonzero cells per condition.  Skipping these genes.
# Clustering observed expression data for each gene
# Setting up parallel back-end using 6 cores
# Notice: Number of permutations is set to zero; using 
#             Kolmogorov-Smirnov to test for differences in distributions
#             instead of the Bayes Factor permutation test
# Classifying significant genes into patterns


```

### MAST

```{r, MAST}
g_data_sub=readRDS("m_g_nofilter_subdata.rds")
groups=c(rep(1,50),rep(2,50))
names(groups)=colnames(g_cdata_sub)
compute_MASTtpmde(g_data_sub,groups,"mouseplate_masttpm.rds")


g_data_sub=readRDS("h_g_nofilter_subdata.rds")
groups=c(rep(1,26),rep(2,26))
names(groups)=colnames(g_data_sub)
compute_MASTtpmde(g_data_sub,groups,"humanplate_masttpm.rds")
```

## Covariate Diagnostics

We could use a few different covariates: strength of the signal defined as the median gene expression unit for the gene, detection rate or a combination of the two. 

```{r,cov,eval=TRUE}
sourceDirectory("/Users/subraman/Documents/research/fdr_jc/src/benchmark-fdr/datasets/R")
source("/Users/subraman/Documents/research/fdr_jc/src/benchmark-fdr/datasets/R/strat_hist.R")

h_cov=compute_covariates_scrnaseq("h_g_nofilter_subdata.rds")

wilcox=readRDS("h_data_wilcox_p.rds")
h_wilcox_cov=data.frame(genes=names(wilcox),pval=wilcox)
h_wilcox_cov=h_wilcox_cov%>%left_join(h_cov,by=c("genes"))

plot_cov_diagnosis(h_wilcox_cov,"humanplate_wilcox")
plot_cov_diagnosis(h_wilcox_cov%>%subset(detection_rate>=0.25),"humanplate_wilcox_filter")

hmast=readRDS("humanplate_masttpm.rds")
h_mast_cov=data.frame(genes=rownames(hmast),pval=hmast$pval)
h_mast_cov=h_mast_cov%>%left_join(h_cov,by=c("genes"))
plot_cov_diagnosis(h_mast_cov,"humanplate_mast")
plot_cov_diagnosis(h_mast_cov%>%subset(detection_rate>=0.25),"humanplate_mast_filter")

h_scdd=readRDS("humanplate_scDatEx_ks.rds")
h_scdd_cov=data.frame(genes=rownames(metadata(h_scdd)$Genes),pval=metadata(h_scdd)$Genes$nonzero.pvalue)
h_scdd_cov$genes=h_cov$genes #scdd changes gene names by appending numeric indices 
h_scdd_cov=h_scdd_cov%>%subset(!is.na(pval))
h_scdd_cov=h_scdd_cov%>%left_join(h_cov,by=c("genes"))
plot_cov_diagnosis(h_scdd_cov,"humanplate_scdd")
plot_cov_diagnosis(h_scdd_cov%>%subset(detection_rate>=0.25),"humanplate_scdd_filter")

m_cov=compute_covariates_scrnaseq("m_g_nofilter_subdata.rds")

wilcox=readRDS("./mouse_plate/m_data_wilcox_p.rds")
m_wilcox_cov=data.frame(genes=names(wilcox),pval=wilcox)
m_wilcox_cov=m_wilcox_cov%>%left_join(m_cov,by=c("genes"))

plot_cov_diagnosis(m_wilcox_cov,"mouseplate_wilcox")

m_wilcox_cov=m_wilcox_cov%>%subset(detection_rate>=0.25)
plot_cov_diagnosis(m_wilcox_cov,"mouseplate_wilcox_filter")

mmast=readRDS("./mouse_plate/mouseplate_masttpm.rds")
m_mast_cov=data.frame(genes=rownames(mmast),pval=mmast$pval)
m_mast_cov=m_mast_cov%>%left_join(m_cov,by=c("genes"))
plot_cov_diagnosis(m_mast_cov,"mouseplate_mast")

m_mast_cov=m_mast_cov%>%subset(detection_rate>=0.25)
plot_cov_diagnosis(m_mast_cov,"mouseplate_mast_filter")


m_scdd=readRDS("mouseplate_scDatEx_ks.rds")
m_scdd_cov=data.frame(genes=rownames(metadata(m_scdd)$Genes),pval=metadata(m_scdd)$Genes$nonzero.pvalue)
m_scdd_cov$genes=m_cov$genes
m_scdd_cov=m_scdd_cov%>%subset(!is.na(pval))
m_scdd_cov=m_scdd_cov%>%left_join(m_cov,by=c("genes"))
plot_cov_diagnosis(m_scdd_cov,"mouseplate_scdd")
plot_cov_diagnosis(m_scdd_cov%>%subset(detection_rate>=0.25),"mouseplate_scdd_filter")

```


