---
title: "Case Study: scRNA-seq (Human data, scDD method)"
author: "Ayshwarya Subramanian and Keegan Korthauer"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Summary

For the application of differential gene expression, we look at scRNAseq and bulk RNAseq data. Here we discuss the analyses of scRNAseq.

```{r echo=FALSE, results='asis'}
library(knitr)
df=data.frame(datasets=c("GSE84465","GSE94383"),
              organism=c("human","mouse"),
              technology=c("smart-seq","smart-seq2"),
              comparison=c("Tumor core vs Periphery",
                           "LPS vs No stimulation"))
kable(df,caption="scRNAseq Datasets")
df1=data.frame(DE_method=c("scDD","MAST"),
               input=c("count","tpm"),
               output=c("p-val","p-val"))
kable(df1,caption="DE methods")
```

We selected two datasets, one each from mouse and human. 
These datasets come from the `Conquer` (Soneson et al., 2018) scRNAseq data collection. `Conquer` provides uniformly processed gene expression data summarized to both count (aggregated from transcript members), and length-scaled TPMs at the gene level. The authors used `salmon` to quantify raw reads and `tximport` to summarize to gene level. The data comes from Smartseq, a plate based sequencing method which results in deeper sequencing of reads at the cost of numbers of cells. 

Here we are interested in performing differential expression analyses to determine
which genes have different expression among the two biological groups in each comparison. 
We also apply two different analysis methods: MAST and scDD. In this vignette, we will
explore the human dataset analysed with scDD.

# Workspace setup

The downloaded plate data are stored in R objects called `MultiAssayExperiment`; this package requires BioConductor release 3.5 and R version 3.4 or greater for installation. 

Note that this Rmd also relies on scDD version 1.3.4 or higher (install from 
Bioconductor devel, or from the master repo on GitHub kdkorthauer/scDD), since 
additional functionality to skip computation of the classification of genes into 
patterns was implemented.

```{r, workspace-setup, results='hide', message=FALSE, warning=FALSE}
library(MultiAssayExperiment)
library(SingleCellExperiment)
library(dplyr)
library(ggplot2)
library(cowplot)
library(impute)
library(scDD)
library(EBSeq)
library(BiocParallel)

if( !packageVersion("scDD") >= '1.3.4'){
  stop("This code relies on scDD version 1.3.4 or higher - see note above")
}

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

# data and results directories
datdir <- "data"
resdir <- "results"
dir.create(datdir, showWarnings = FALSE)
dir.create(resdir, showWarnings = FALSE)

# set up parallel backend
cores <- 8
BiocParallel::register(MulticoreParam(workers = cores))
```

# Data preparation

## Data download

We first download the data in the data directory.`conquer` provides the processed datasets in convenient .rds files. First, we download the Human dataset:
GSE84465.

```{r, humandata-download}
datfile <- file.path(datdir, "GSE84465.rds")

if (!file.exists(datfile)){
  download.file("http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE84465.rds",
    destfile=datfile)
}
```

## Data processing and filtering

For data processing, we follow practices and recommendations from Soneson et al., 2017. Briefly, we limit ourselves to 2-group comparisons. We do not filter genes as filtering may affect the subsequent false discovery control steps. We only filter out genes that are expressed in 
fewer than 1% of cells. 
We will filter out cells with extremely low mapping rate
(below 20%), as these represent poor quality cells that may end up having negative
size factor estimates during normalization. For the same reason, we also filter cells
which have extremely low proportion of genes detected (less than 5%).

This dataset (from PMC5810554) contains cells from four Glioblastoma patients, 
sampled from both the tumor core as well as the tissue peripheral to the tumor. 
We'll compare the Neoplastic cells in the tumor core vs periphery. We'll exclude 
the two samples where only a few neoplastic cells were found in the periphery.

```{r}
datfile <- file.path(datdir, "human_GSE84465_scdd.rds")

if (!file.exists(datfile)){
  
  # load MultiAssayExperiment
  humandata <- readRDS(file.path(datdir, "GSE84465.rds"))
  
  # exclude cells with extremely low mapping or feature rate (poor quality cells)
  lowMap <- which(metadata(humandata)$salmon_summary$percent_mapped < 20 |
                  colMeans(assay(experiments(humandata)[[1]], "count") > 0) < 0.05)
  if (length(lowMap) > 0){
    humandata <- humandata[, -lowMap]
    message("Removed ", length(lowMap), " cells with mapping rate below 20%",
            " or fewer than 5% of genes detected.")
  }
  
  # subset by groups of interest
  humandata <- humandata[, colData(humandata)$patient.id.ch1 %in% 
                           c(" BT_S4", " BT_S2") &
                           colData(humandata)$cell.type.ch1 == " Neoplastic" ]
  cdata=colData(humandata)
  cdata$group <- cdata$tissue.ch1

  # subset by gene expression TPM & count assays
  humandata <- experiments(humandata)[[1]]
  assays(humandata) <- assays(humandata)[c("TPM", "count")]
  colData(humandata) <- cdata

  # subset by genes with detection rate > 5% in at least one condition
  levs <- unique(colData(humandata)$group)
  humandata <- humandata[rowMeans(assay(humandata, "count")[, 
                            colData(humandata)$group == levs[1]] > 0) > 0.05 |
                         rowMeans(assay(humandata, "count")[, 
                            colData(humandata)$group == levs[2]] > 0) > 0.05, ]
  
  # remove spike-in controls
  humandata <- humandata[!grepl("ERCC-", rownames(humandata)),]
 
  # save SummarizedExperiment 
  saveRDS(humandata, datfile)
}else{
  humandata <- readRDS(datfile)
}

# look at number of cells in each group
table(colData(humandata)$group)
```

For the human dataset, there are a total of `r ncol(humandata)` cells and `r nrow(humandata)` genes. 

# Data analysis

## Differential testing

We perform differential expression testing using scDD which implicitly models any heterogeneity in expression. 

We apply scDD to normalized log-transformed counts. 
The p-value results are added to the `rowData` slot of the `SummarizedExperiment` 
data objects to ensure that gene metadata stays intact.

```{r, scdd}
# function to add scDD p-values
compute_scdd_sc <- function(dat){
  # check if already computed
  if (! "scdd_p" %in% colnames(rowData(dat))){
    sce <- SingleCellExperiment(assays=list(count=assay(dat, "count")),
                                colData=data.frame(colData(dat)))
    # scran normalization
    libsize <- scran::computeSumFactors(assay(sce, "count"))
    normcounts(sce) <- GetNormalizedMat(assay(sce, "count"), libsize)

    res <- results(scDD(sce, 
             condition = "group",
             prior_param = list(alpha=0.01, mu0=0, s0=0.01, a0=0.01, b0=0.01), 
             testZeroes = TRUE,
             categorize = FALSE))
    rowData(dat)$scdd_p_nonzero <- res$nonzero.pvalue
    rowData(dat)$scdd_p_zero <- res$zero.pvalue
    rowData(dat)$scdd_p_combined <- res$combined.pvalue
  }else{
     message("scDD p-values already computed.")
  }
  return(dat)
}

humandata <- compute_scdd_sc(humandata)

# save results
saveRDS(humandata, datfile)
```

## Covariate Diagnostics

In scrnaseq data, strength of the signal can be a affected by the of level of gene expression. 
We will explore two potential covariates related to expression level: mean nonzero expression 
and detection rate (defined as the proportion of cells expressing the gene at a nonzero level). 
These will also be added to the `rowData` slot of the `SummarizedExperiments`.

```{r,cov}
add_covariates_scrnaseq <- function(dat){
  rowData(dat)$meanExp <- apply(assay(dat, "count"), 1, 
                                function(x) mean(x[x>0]))
  rowData(dat)$detection <- rowMeans(assay(dat, "count") > 0)
  return(dat)
}

humandata <- add_covariates_scrnaseq(humandata)

# save results
saveRDS(humandata, datfile)
```

### Covariate one: Mean Nonzero Expression

For each covariate, we'll examine the covariate diagnostic plots.

```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
rank_scatter(data.frame(rowData(humandata)), 
             pvalue="scdd_p_combined", covariate="meanExp") + 
  ggtitle("scDD, Covariate 1: Mean Nonzero Expression")
```
 
```{r, fig.width=10, fig.height=3.2, message=FALSE}
strat_hist(data.frame(rowData(humandata)), 
           pvalue="scdd_p_combined", covariate="meanExp", maxy=12,
           main = "scDD, Covariate 1: Mean Nonzero Expression") 
```

The mean nonzero expression appears to be informative and approximately satisfies the 
assumptions 

### Covariate two: Detection Rate

Next we look at the detection rate covariate.

```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
rank_scatter(data.frame(rowData(humandata)), 
             pvalue="scdd_p_combined", covariate="detection") + 
  ggtitle("scDD, Covariate 2: Detection Rate")
```
 
```{r, fig.width=10, fig.height=3.2, message=FALSE}
strat_hist(data.frame(rowData(humandata)), 
           pvalue="scdd_p_combined", covariate="detection", maxy=12,
           main = "scDD, Covariate 2: Detection Rate") 
```
## Multiple-Testing Correction

First, we'll create an object of `BenchDesign` class to hold the data and 
add the benchmark methods to the `BenchDesign` object.

```{r, benchdesign, message=FALSE}
bd <- initializeBenchDesign()
```

Now, we're ready to construct the `SummarizedBenchmark` object, which will run
the functions specified in each method (these are actually sourced in from the
                                        helper scripts). 
                                        
### Covariate one: Mean expression

First we'll include the mean nonzero expression covariate.

```{r, sb1, results="hide", message=FALSE}
res1 <- file.path(resdir, "sb_GSE84465_scdd_mean.rds")

if (!file.exists(res1)){
  sb1 <- bd %>% buildBench(data.frame(rowData(humandata)) %>% 
                                   na.omit() %>%
                                   mutate(pval=scdd_p_combined,
                                          ind_covariate=meanExp), 
                          ftCols = c("meanExp"),
                          parallel=TRUE)

  saveRDS(sb1, file=res1)
}else{
  sb1 <- readRDS(res1)
}
```

### Covariate two: Detection Rate

Now, we'll repeat the multiple testing correction using the other covariate (detection rate):

```{r, sb2, results="hide", message=FALSE}
res2 <- file.path(resdir, "sb_GSE84465_scdd_det.rds")

if (!file.exists(res2)){
  sb2 <- bd %>% buildBench(data.frame(rowData(humandata)) %>% 
                                   na.omit() %>%
                                   mutate(pval=scdd_p_combined,
                                          ind_covariate=detection), 
                          ftCols = c("detection"),
                          parallel=TRUE)

  saveRDS(sb2, file=res2)
}else{
  sb2 <- readRDS(res2)
}
```

## Benchmark Metrics

Next, we'll add the default performance metric for q-value assays and 
plot the results. We'll start with covariate one.

### Covariate one: Mean Nonzero Expression

First, we have
to rename the assay to 'qvalue'.

```{r, metrics1}
# rename assay to qvalue
assayNames(sb1) <- "qvalue"
sb1 <- addDefaultMetrics(sb1)
```

Now, we'll plot the results.

```{r, plot1, results="hide"}
# plot nrejects by method overall and stratified by covariate
rejections_scatter(sb1,
                   supplementary=FALSE) +
  ggtitle("scDD, Covariate 1: Mean Nonzero Expression")

rejection_scatter_bins(sb1, covariate="meanExp", bins=4,
                       supplementary=FALSE) +
  ggtitle("scDD, Covariate 1: Mean Nonzero Expression")

# upset plot 
plotFDRMethodsOverlap(sb1, 
                      alpha=0.05, nsets=ncol(sb1),
                      order.by="freq", decreasing=TRUE,
                      supplementary=FALSE) 
```

```{r, fig.width=8, fig.height=3.5}
mcols(sb1)$ind_covariate <- mcols(sb1)$meanExp
covariateHeatmap(sb1, alpha=0.05, covname="ind_covariate", nbins=25, 
                 linePlot = TRUE) +
  ggtitle("scDD, Covariate 1: Mean Nonzero Expression")
```

### Covariate two: Detection Rate

Next, we'll look at the performance metrics for the other covariate (detection rate).

```{r, metrics2}
# rename assay to qvalue
assayNames(sb2) <- "qvalue"
sb2 <- addDefaultMetrics(sb2)
```

Now, we'll plot the results.

```{r, plot2, results="hide"}
# plot nrejects by method overall and stratified by covariate
rejections_scatter(sb2, supplementary=FALSE) +
  ggtitle("scDD, Covariate 2: Detection Rate")

rejection_scatter_bins(sb2, covariate="detection", bins=4,
                       supplementary=FALSE) +
  ggtitle("scDD, Covariate 2: Detection Rate")

# upset plot 
plotFDRMethodsOverlap(sb2, 
                      alpha=0.05, nsets=ncol(sb2),
                      order.by="freq", decreasing=TRUE,
                      supplementary=FALSE)
```

```{r, fig.width=8, fig.height=3.5}
mcols(sb2)$ind_covariate <- mcols(sb2)$detection
covariateHeatmap(sb2, alpha=0.05, covname="ind_covariate", nbins=25, 
                 linePlot = TRUE) +
  ggtitle("scDD, Covariate 2: Detection Rate")
```

# Session Information

```{r}
sessionInfo()
```

