---
title: "Simulation Study: AdaPT with null model option (step covariate)" 
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

# Summary

In this set of simulations, we consider settings with both null and non-null
tests with an informative covariate. The covariate is sampled uniformly from
the interval [0, 1], and the conditional probability of a test being non-null
is a non-smooth (monotone) step function of the covariate. We include simulation
results again with Gaussian, t-distributed, and Chi-Squared distributed test statistics.

This set of simulations differs from `simulations-informative-step.Rmd` only
in the implementation of the AdaPT method for multiple testing correction.
In the primary set of simulation results (`simulations-informative-step.Rmd`),
we observed that the AdaPT method can lose power when the informative covariate
is only weakly informative. While the method still controls the FDR, we observe
that the loss in power is substantial, with AdaPT achieving a lower mean TPR
across 100 replications when using a weakly informative covariate rather than a
completely uninformative (random) covariate.

In this set of simulations, we assess whether this loss of power with a weakly
informative covariate can be mitigated by allowing the AdaPT model selection
procedure to decide between the same collection of spline models used before, and
an intercept-only (null) model, which makes no use of the covariate. This approach
was suggested by the authors of the AdaPT method through personal communication. 

Two additional implementations of the AdaPT model are included in this analysis:
`adapt-withnull`, the approach described above, and `adapt-onlynull`, where the method
is forced to ignore the informative covariate. The latter (`onlynull`) implementation
provides a baseline for comparing the two other AdaPT parameter settings.

# Workspace Setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(SummarizedBenchmark)
library(parallel)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
resdir <- "results"
dir.create(resdir, showWarnings = FALSE, recursive = TRUE)

## intermediary files we create below
gauss_file <- file.path(resdir, "nullAdaPT-informative-step-benchmark-gaussian.rds")
tdist_file <- file.path(resdir, "nullAdaPT-informative-step-benchmark-t5.rds")
tdist11_file <- file.path(resdir, "nullAdaPT-informative-step-benchmark-t11.rds")
chisq_file <- file.path(resdir, "nullAdaPT-informative-step-benchmark-chisq4.rds")

## number of cores for parallelization
cores <- 20
B <- 100

## define bechmarking design
bd <- initializeBenchDesign()
```

As described in `simulations-null.Rmd`, we include Scott's FDR Regression in the analysis
for simulations with Gaussian or t-distributed test statistics. Again, we include both
`nulltype = "empirical"` and `nulltype = "theoretical"`. 

```{r}
bdplus <- bd
bdplus <- addBMethod(bdplus, "fdrreg-t",
                     FDRreg::FDRreg,
                     function(x) { x$FDR },
                     z = test_statistic,
                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
                     nulltype = 'theoretical',
                     control = list(lambda = 0.01))
bdplus <- addBMethod(bdplus, "fdrreg-e",
                     FDRreg::FDRreg,
                     function(x) { x$FDR },
                     z = test_statistic,
                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
                     nulltype = 'empirical',
                     control = list(lambda = 0.01))
```

We add a couple alternative calls to AdaPT to test whether the method performance improves in cases
of low "informativeness" if given the option to not use the informative covariate. This option is
coded by the null model, `"pi0 ~ 1"`, where only an intercept term is included in the model. Specifically,
we add two alternative calls to the benchmark - one call with the null model added as one of the candidate
models to `pi_formulas=` and `mu_formulas=`, and a second call with the null model given as the only
candidate model to `pi_formulas=` and `mu_formulas=`.

```{r}
bdplus <- addBMethod(bdplus, "adapt-withnull",
                     adaptMT::adapt_glm,
                     function(x) { x$qvals },
                     pvals = pval,
                     x = data.frame(icov = ind_covariate),
                     pi_formulas = c("~ 1", paste0("splines::ns(icov, df = ", seq(2, 10, 2), ")")),
                     mu_formulas = c("~ 1", paste0("splines::ns(icov, df = ", seq(2, 10, 2), ")")),
                     alphas = 0)

bdplus <- addBMethod(bdplus, "adapt-onlynull",
                     adaptMT::adapt_glm,
                     function(x) { x$qvals },
                     pvals = pval,
                     x = data.frame(icov = ind_covariate),
                     pi_formulas = "~ 1",
                     mu_formulas = "~ 1",
                     alphas = 0)

candycols2 <- rbind(candycols,
                    data.frame(Method = c("adapt-withnull", "adapt-onlynull"),
                               col = c("darkgoldenrod3", "darkgoldenrod3"),
                               lty = c("dashed", "dotted"),
                               stringsAsFactors = FALSE))
```

All simulation settings will share the following parameters.

```{r parameters-shared}
m <- 20000                        # integer: number of hypothesis tests
pi0 <- pi0_step(0.90)             # numeric: proportion of null hypotheses
icovariate <- runif               # functional: independent covariate
```

Simulation results will be presented excluding a subset of methods, and
for certain plots (upset plots), a single alpha cutoff will be used.

```{r}
excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
ualpha <- 0.05
```

# Gaussian Setting

First, we consider the setting with Gaussian test statistics.

## Data Simulation

```{r gauss-parameters}
es_dist <- rnorm_generator(3)       # functional: dist of alternative test stats
ts_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
null_dist <- rnorm_2pvaluer(1)    # functional: dist to calc p-values
seed <- 608
```

We next run the simulations (including Scott's FDR Regression).

```{r gauss-run-simulation}
if (file.exists(gauss_file)) {
    res <- readRDS(gauss_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
                    pi0 = pi0, es_dist = es_dist, icovariate = icovariate,
                    ts_dist = ts_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = gauss_file)
}
res_i <- lapply(res, `[[`, "informative")
res_u <- lapply(res, `[[`, "uninformative")
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r gauss-one-simulation}
onerun <- simIteration(1, bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, gauss-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, gauss-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r gauss-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)
```

Since we are primarily interested in comparing just the AdaPT methods, we plot these
metrics excluding all other methods.

```{r gauss-metrics-averages-adapt, results = "hide"}
plotsim_average(resdf, met="rejections",
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r gauss-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size",
                  palette = candycols2)

covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate",
                  palette = candycols2)
```

We also look at the FDR as a function of the independent covariate.

```{r gauss-metrics-covlineplotFDR, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR",
                  palette = candycols2)
```

We also compare the simulation results with and without an informative covariate. 

```{r gauss-metrics-differences, results = "hide"}
resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))

resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            select(resdfu, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            by = c("rep", "blabel", "param.alpha", "key",
                                   "performanceMetric", "alpha"),
                            suffix = c(".info", ".uninfo"))
resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)

plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)
```

Again, we recreate the plots after subsetting on AdaPT methods. 

```{r gauss-metrics-differences-adapt, results = "hide"}
plotsim_average(resdfiu, met="rejections", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)
```

# Student's t Setting (df = 5)

Next, we consider the setting with t-distributed test statistics.

## Data Simulation

```{r t5-parameters}
es_dist <- rnorm_generator(6)    # functional: dist of alternative test stats
ts_dist <- rt_perturber(5)  # functional: sampling dist/noise for test stats
null_dist <- rt_2pvaluer(5)    # functional: dist to calc p-values
seed <- 815
```

For the t-distributed setting, we must specify the number of degrees of freedom
for ASH. We add an additional parameter to the `ashq` method with the corresponding
degrees of freedom of the test statistic distribution.

```{r}
bdplust <- modifyBMethod(bdplus, "ashq", df = 5)
```

We next run the simulations (including Scott's FDR Regression and ASH with degrees of
freedom specified).

```{r t5-run-simulation}
if (file.exists(tdist_file)) {
    res <- readRDS(tdist_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplust, m = m,
                    pi0 = pi0, es_dist = es_dist, icovariate = icovariate,
                    ts_dist = ts_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = tdist_file)
}
res_i <- lapply(res, `[[`, "informative")
res_u <- lapply(res, `[[`, "uninformative")
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r t5-one-simulation}
onerun <- simIteration(1, bdplust, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, t5-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, t5-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r t5-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)
```

Since we are primarily interested in comparing just the AdaPT methods, we plot these
metrics excluding all other methods.

```{r t5-metrics-averages-adapt, results = "hide"}
plotsim_average(resdf, met="rejections",
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r t5-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size",
                  palette = candycols2)

covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate",
                  palette = candycols2)
```

We also look at the FDR as a function of the independent covariate.

```{r t5-metrics-covlineplotFDR, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR",
                  palette = candycols2)
```

We also compare the simulation results with and without an informative covariate. 

```{r t5-metrics-differences, results = "hide"}
resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))

resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            select(resdfu, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            by = c("rep", "blabel", "param.alpha", "key",
                                   "performanceMetric", "alpha"),
                            suffix = c(".info", ".uninfo"))
resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)

plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)
```

Again, we recreate the plots after subsetting on AdaPT methods. 

```{r t5-metrics-differences-adapt, results = "hide"}
plotsim_average(resdfiu, met="rejections", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)
```

# Student's t Setting (df = 11)

Next, we consider a second setting with t-distributed test statistics.

## Data Simulation

```{r t11-parameters}
es_dist <- rnorm_generator(3)     # functional: dist of alternative test stats
ts_dist <- rt_perturber(11)  # functional: sampling dist/noise for test stats
null_dist <- rt_2pvaluer(11)    # functional: dist to calc p-values
seed <- 9158
```

For the t-distributed setting, we must specify the number of degrees of freedom
for ASH. We add an additional parameter to the `ashq` method with the corresponding
degrees of freedom of the test statistic distribution.

```{r}
bdplust <- modifyBMethod(bdplus, "ashq", df = 11)
```

We next run the simulations (including Scott's FDR Regression and ASH with degrees of
freedom specified).

```{r t11-run-simulation}
if (file.exists(tdist11_file)) {
    res <- readRDS(tdist11_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplust, m = m,
                    pi0 = pi0, es_dist = es_dist, icovariate = icovariate,
                    ts_dist = ts_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = tdist11_file)
}
res_i <- lapply(res, `[[`, "informative")
res_u <- lapply(res, `[[`, "uninformative")
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r t11-one-simulation}
onerun <- simIteration(1, bdplust, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, t11-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, t11-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r t11-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)
```

Since we are primarily interested in comparing just the AdaPT methods, we plot these
metrics excluding all other methods.

```{r t11-metrics-averages-adapt, results = "hide"}
plotsim_average(resdf, met="rejections",
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r t11-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size",
                  palette = candycols2)

covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate",
                  palette = candycols2)
```

We also look at the FDR as a function of the independent covariate.

```{r t11-metrics-covlineplotFDR, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR",
                  palette = candycols2)
```

We also compare the simulation results with and without an informative covariate. 

```{r t11-metrics-differences, results = "hide"}
resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))

resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            select(resdfu, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            by = c("rep", "blabel", "param.alpha", "key",
                                   "performanceMetric", "alpha"),
                            suffix = c(".info", ".uninfo"))
resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)

plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)
```

Again, we recreate the plots after subsetting on AdaPT methods. 

```{r t11-metrics-differences-adapt, results = "hide"}
plotsim_average(resdfiu, met="rejections", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)
```

# Chi-Squared Setting

Finally, we consider the setting with chi-squared distributed test statistics.

## Data Simulation

```{r chisq4-parameters}
es_dist <- rnorm_generator(15)       # functional: dist of alternative test stats
ts_dist <- rchisq_perturber(4)  # functional: sampling dist/noise for test stats
null_dist <- rchisq_pvaluer(4)     # functional: dist to calc p-values
seed <- 1023
```

For the chi-squared distributed setting, we must change the "mode" setting for ASH
from the default of 0 to `"estimate"` since the mode of null and alternative test statistics
are no longer centered at 0. While both approximate Normality and unimodality of effects are
violated in this simulation setting, by allowing the mode to be estimated, rather than forced
to 0, should return more comparable results.

```{r}
bdchi <- modifyBMethod(bdplus, "ashq", mode = "empirical")
```

We also drop the FDRreg methods from the chi-squared setting.

```{r}
bdchi <- dropBMethod(bdchi, "fdrreg-t")
bdchi <- dropBMethod(bdchi, "fdrreg-e")
```

We next run the simulations. We do not include FDR Regression because the test statistics
are not approximately normally distributed.

```{r chisq4-run-simulation}
if (file.exists(chisq_file)) {
    res <- readRDS(chisq_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdchi, m = m,
                    pi0 = pi0, es_dist = es_dist, icovariate = icovariate,
                    ts_dist = ts_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = chisq_file)
}
res_i <- lapply(res, `[[`, "informative")
res_u <- lapply(res, `[[`, "uninformative")
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r chisq4-one-simulation}
onerun <- simIteration(1, bdchi, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, chisq4-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, chisq4-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r chisq4-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met = "rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 

plotsim_average(resdf, met = "FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 

plotsim_average(resdf, met = "TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 

plotsim_average(resdf, met = "TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

Since we are primarily interested in comparing just the AdaPT methods, we plot these
metrics excluding all other methods.

```{r chisq4-metrics-averages-adapt, results = "hide"}
plotsim_average(resdf, met="rejections",
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate. Again, we plot these without `ashq`.

```{r chisq4-metrics-covlineplot, results = "hide"}
covariateLinePlot(lapply(res_i, function(x) { x[, -which(colnames(x) == "ashq")] }),
                  alpha = ualpha, covname = "effect_size")

covariateLinePlot(lapply(res_i, function(x) { x[, -which(colnames(x) == "ashq")] }),
                  alpha = ualpha, covname = "ind_covariate")
```

We also look at the FDR as a function of the independent covariate.

```{r chisq4-metrics-covlineplotFDR, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR")
```

We also compare the simulation results with and without an informative covariate. 

```{r chisq4-metrics-differences, results = "hide"}
resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))

resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            select(resdfu, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            by = c("rep", "blabel", "param.alpha", "key",
                                   "performanceMetric", "alpha"),
                            suffix = c(".info", ".uninfo"))
resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)

plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)

plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)

plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)

plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
```

Again, we recreate the plots after subsetting on AdaPT methods. 

```{r chisq4-metrics-differences-adapt, results = "hide"}
plotsim_average(resdfiu, met="rejections", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)
```

# Session Info

```{r}
sessionInfo()
```
