---
title: "Manuscript Figures for Genome Biology (simulation subset)"
author: "Rafalab"
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Running this

```{r load-workspace}
## Load packages and source benchmark FDR
library(tidyverse)
library(cowplot)
library(ggthemes)
library(grid)
library(SummarizedBenchmark)

## load helper functions
for (f in list.files("../datasets/R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

# assumes sb objects are in the following location
resmet_file <- file.path("..", "datasets", "simulations", "results-summary", "result-metrics.rds")

# set up results directory
outdir <- "./figures"
dir.create(outdir, showWarnings = FALSE)

methodset <- c("bonf", "bh", "ihw", "qvalue", "bl-df03", "lfdr",
               "fdrreg-e", "fdrreg-t", "ashq", "adapt-glm")
```

We use the standardize candy color scheme and line types for the plots.

```{r}
col <- as.character(candycols$col)
names(col) <- as.character(candycols$Method)
lty <- as.character(candycols$lty)
names(lty) <- as.character(candycols$Method)
```

To generate the figures in this document, the simulation results must first be aggregated by running the code at `datasets/simulations/simulations-summary.Rmd`.

```{r load-sim-metrics}
resmet <- readRDS(resmet_file)
```

Unfortunately, because the plots in this section illustrate FDR and TPR acorss varying simulation settings rather than varying nominal FDR cutoff, we cannot simply use the `plotsim_average()` function to generate plots. Instead, we define the following helper function to consistently generate plots similar to those output by `plotsim_average()` with arbitrary covariate, e.g. number of tests or proportion of null hypotheses, as the x-axis. The function assumes that the input table only includes FDR or TPR values for each method at a single nominal alpha cutoff. In these figures, we only plot FDR and TPR values at the nominal alpha cutoff of 0.05.

```{r genplot-function}
genplot <- function(tab, cov, xt = "", met = c("FDR", "TPR"), ebw = 0.0025) {
    met <- match.arg(met)
    cov <- rlang::enquo(cov)
    gp <- dplyr::filter(tab, key == met) %>%
        ggplot(aes(x = !!cov, y = mean.info, color = Method, group = Method)) +
        geom_line(aes(linetype = Method), alpha = 0.85) +
        geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = ebw, alpha=0.5) +
        scale_linetype_manual(values = lty) + 
        scale_color_manual(values = col) +
        scale_y_continuous(met, labels = scales::percent) +
        expand_limits(y = 0) +
        theme_classic() +
        theme(axis.title = element_text(face = "bold", size = 10),
              plot.title = element_text(face = "bold", size = 14))
    if (met == "FDR") {
        if (rlang::quo_name(cov) == "alpha") {
            gp <- gp + geom_abline(lty = 2, color = "blue", alpha = 1/2) 
        } else {
            gp <- gp + geom_hline(yintercept = 0.05, lty = 2, color = "blue", alpha = 1/2) 
        }
    }
    gp
}
```

# Figure 3

This is a faceted figure with results from the pure simulation studies showing several key results from a subset of simulation settings. Namely, the figure includes plots of FDR and TPR values at a nominal FDR cutoff of 0.05 over 100 replications illustrating the impact of (A,B) changing total number of tests, (C,D) changing the proportion of non-null tests, and (E,F) changing the informativeness of a covariate. 

```{r Figure3, fig.width = 10, fig.height = 12}
## Figure 3A,B: FDR/TPR of methods across varying number of tests at alpha = 0.05
res_ntests <- dplyr::filter(resmet, setting == "varyingntests", alpha == 0.05)
p3a <- genplot(res_ntests, cov = ntests, met = "FDR", ebw = 0.05) +
    scale_x_continuous("Number of Tests", trans = "log10",
                       breaks = c(1e2, 5e2, 1e3, 5e3, 1e4, 5e4))
p3b <- genplot(res_ntests, cov = ntests, met = "TPR", ebw = 0.05) +
    scale_x_continuous("Number of Tests", trans = "log10",
                       breaks = c(1e2, 5e2, 1e3, 5e3, 1e4, 5e4))

## add title to 3A,B
p3ab <- plot_grid(p3a + guides(color = FALSE, linetype = FALSE),
                  p3b + guides(color = FALSE, linetype = FALSE),
                  labels = LETTERS[1:2], ncol = 2)
p3abTitle <- ggdraw() +
    draw_label("Simulation performance across varying numbers of hypotheses",
               fontface = 'bold')
p3ab <- plot_grid(p3abTitle, p3ab, ncol = 1, rel_heights = c(0.1, 1))

## Figure 3C,D: FDR/TPR of methods across varying null proportion at alpha = 0.05
res_pi0 <- dplyr::filter(resmet, setting == "varyingpi0", alpha == 0.05)
p3c <- genplot(res_pi0, cov = 100-pi0, met = "FDR", ebw = 2.5) +
    scale_x_continuous("Proportion of Non-Null Hypotheses",
                       breaks = 100 - c(5, seq(10, 90, by = 10), 95, 99))
p3d <- genplot(res_pi0, cov = 100-pi0, met = "TPR", ebw = 2.5) +
    scale_x_continuous("Proportion of Non-Null Hypotheses",
                       breaks = 100 - c(5, seq(10, 90, by = 10), 95, 99))

## add title to 3C,D
p3cd <- plot_grid(p3c + guides(color = FALSE, linetype = FALSE),
                  p3d + guides(color = FALSE, linetype = FALSE),
                  labels = LETTERS[3:4], ncol = 2)
p3cdTitle <- ggdraw() +
    draw_label("Simulation performance across varying non-null proportion",
               fontface = 'bold')
p3cd <- plot_grid(p3cdTitle, p3cd, ncol = 1, rel_heights = c(0.1, 1))

## Figure 3E,F: FDR/TPR of methods across varying informativeness at alpha = 0.05
res_info <- dplyr::filter(resmet, setting == "varyinginfo", alpha == 0.05)
p3e <- genplot(res_info, cov = inform, met = "FDR", ebw = 2.5) +
    scale_x_continuous("Informativeness", breaks = seq(0, 100, by = 20))
p3f <- genplot(res_info, cov = inform, met = "TPR", ebw = 2.5) +
    scale_x_continuous("Informativeness", breaks = seq(0, 100, by = 20))

## add title to 3E,F
p3ef <- plot_grid(p3e + guides(color = FALSE, linetype = FALSE),
                  p3f + guides(color = FALSE, linetype = FALSE),
                  labels = LETTERS[5:6], ncol = 2)
p3efTitle <- ggdraw() +
    draw_label("Simulation performance across varying informativeness",
               fontface = 'bold')
p3ef <- plot_grid(p3efTitle, p3ef, ncol = 1, rel_heights = c(0.1, 1))

## pull Figure 3 together
Fig3 <- plot_grid(p3ab, p3cd, p3ef, ncol = 1)
Fig3 <- plot_grid(Fig3, get_legend(p3a), rel_widths = c(1, .2))
Fig3
ggsave(file.path(outdir, "Figure3.pdf"), width=10, height=12)
```

# Figure 3+

This is a faceted figure with results from the pure simulation studies showing the behavior of methods across different test statistic distributions. Test statistics were simulated from Normal, t with 11 degrees of freedom, t with 5 degrees of freedom, and Chi-squared with 4 degrees of freedom. 

```{r Figure3plus, fig.width = 10, fig.height = 12}
## Figure 3plus A,B: FDR/TPR of methods across varying test statistics distribution
res_noise <- dplyr::filter(resmet, setting == "informative-cubic")
levels(res_noise$dist) <- gsub("Noise:", "Distribution:", levels(res_noise$dist))
p3pa <- genplot(res_noise, cov = alpha, met = "FDR") +
    facet_grid(. ~ dist, labeller = label_parsed)
p3pb <- genplot(res_noise, cov = alpha, met = "TPR") +
    facet_grid(. ~ dist, labeller = label_parsed)

## create title for 3plus
p3pTitle <- ggdraw() +
    draw_label("Simulation performance across test statistic distributions",
               fontface = 'bold', size = 20)

## pull Figure 3plus together
Fig3p <- plot_grid(p3pa + guides(color = FALSE, linetype = FALSE),
                   p3pb + guides(color = FALSE, linetype = FALSE),
                   labels = LETTERS[1:2], ncol = 1)
Fig3p <- plot_grid(p3pTitle, Fig3p, ncol = 1, rel_heights = c(0.1, 1))
Fig3p <- plot_grid(Fig3p, get_legend(p3pa), rel_widths = c(7.5, 1))
Fig3p
ggsave(file.path(outdir, "Figure3plus.pdf"), width=14, height=8)
```

# Session information

```{r}
sessionInfo()
```
