---
title: "Simulation Figures for Genome Biology (simulations)"
author: "Rafalab"
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Figure2-load-data}
# Load packages and source benchmark FDR
library(tidyr)
library(dplyr)
library(ggplot2)
library(magrittr)
library(cowplot)
library(tibble)
library(ggthemes)
library(grid)
library(SummarizedBenchmark)

## load helper functions
for (f in list.files("../datasets/R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

# assumes sb objects are in the following location
path <- file.path("..", "datasets", "simulations", "results")

# set up results directory
outdir <- "./figures"
dir.create(outdir, showWarnings = FALSE)

# set alpha cutoff for plots that are fixed at a certain value of alpha
alpha <- 0.10

methodset <- c("bonf", "bh", paste0("ihw-a", round(alpha * 100)),
               "qvalue", "bl-df03", "lfdr",
               "scott-empirical", "scott-theoretical", "ashq")
```

# Data Preparation

First, we load summaries for all simulations results.

```{r}
summary_file <- paste0("allres_alpha", 100*alpha, ".rds")

if (file.exists(summary_file)) {
    res <- readRDS(summary_file)
} else {
    res_files <- list.files(path, "\\.rds$", full.names = TRUE)
    res <- lapply(res_files, function(x) {
        zz <- readRDS(x)
        zz <- plotsim_standardize(zz, alpha = alpha)
        zz <- dplyr::select(zz, rep, blabel, key, alpha, value)

        zz <- dplyr::group_by(zz, blabel, key, alpha)
        zz <- dplyr::summarize(zz,
                               se = sd(value, na.rm = TRUE) / sqrt(sum(!is.na(value))),
                               q10 = quantile(value, .1, na.rm = TRUE),
                               q90 = quantile(value, .9, na.rm = TRUE),
                               med_value = median(value, na.rm = TRUE),
                               mean_value = mean(value, na.rm = TRUE),
                               nNA = sum(is.na(value)))
        dplyr::ungroup(zz)
    })
    names(res) <- gsub("\\.rds$", "", basename(res_files))
    saveRDS(res, summary_file)
}

## merge all simulation summaries
res <- bind_rows(res, .id = "setting")

## create shorter simulation setting labels
res <- dplyr::mutate(res, dist = gsub(".*?-benchmark-(.*)", "\\1", setting),
                     setting = gsub("(.*?)-benchmark-.*", "\\1", setting))

## only use subset of methods for plotting
res <- dplyr::filter(res, blabel %in% methodset)
res <- dplyr::rename(res, Method = blabel)

## remove suffixes for unused tuning parameters
res <- dplyr::mutate(res, Method = gsub("-df03", "", Method),
                     Method = gsub("(-a)(.*)", "", Method))
methodset <- gsub("-df03", "", methodset)
methodset <- gsub("(-a)(.*)", "", methodset)

## add candy color scheme from datasets/R/colorPalette.R
res <- dplyr::left_join(res, candycols, by = "Method")

## treat methods as factors for forced ordering
res <- dplyr::mutate(res, Method = factor(Method, levels = methodset))
```

# Null Simulations

We subset on simulation results that look at informative and non-informative covariates across
different noise distributions.

```{r}
res_nullset <- dplyr::filter(res, setting == "null")
res_nullset <- dplyr::mutate(res_nullset, dist = factor(dist, levels = c("gaussian", "t5", "chisq4")))
res_nullset <- tidyr::complete(res_nullset, setting, Method, dist, key)

col <- as.character(res_nullset$col)
names(col) <- as.character(res_nullset$Method)
```

For null simulations, rather than the false discovery rate, we take a look at the number
false positives since the FDR will be 1 if any rejections are made.

In these plots, `ashq`, `scott-theoretical` and `scott-theoretical` are excluded for the Chi-Squared
simulation settings.

```{r, fig.width = 8, fig.height = 5}
dplyr::filter(res_nullset, key == "rejectprop") %>%
    filter(! (Method == "ashq" & dist == "chisq4"),
           ! (grepl("scott", Method) & dist == "chisq4")) %>%
    ggplot(aes(x = Method, y = mean_value, color = Method)) +
    geom_point() +
    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se)) + 
    facet_grid(. ~ dist) +
    scale_color_manual(values = col, guide = FALSE) +
    scale_y_continuous("Average % Rejected Tests", breaks = seq(0, 1, .01), labels = scales::percent) +
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
    ggtitle("Rejected proportion in null simulations (average over 100 replications)") 
```

Additionally, we take a look at the FWER (the proportion of simulation replicates out of 100
with at least one rejection).

```{r, fig.width = 8, fig.height = 5}
gp_nullA <- dplyr::filter(res_nullset, key == "FWER") %>%
    filter(! (Method == "ashq" & dist == "chisq4"),
           ! (grepl("scott", Method) & dist == "chisq4")) %>%
    ggplot(aes(x = Method, y = mean_value, color = Method)) +
    geom_point() +
    scale_color_manual(values = col, guide = FALSE) +
    facet_grid(. ~ dist) +
    scale_y_continuous("FWER", breaks = seq(0, 1, .2), labels = scales::percent) +
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
    ggtitle("Proportion of replications with at least one false positive (over 100 replications)")
gp_nullA
```

Finally, we plot the TNR as was done for the yeast in silico experiments/simulations.

```{r, fig.width = 8, fig.height = 5}
gp_nullB <- dplyr::filter(res_nullset, key == "TNR") %>%
    filter(! (Method == "ashq" & dist == "chisq4"),
           ! (grepl("scott", Method) & dist == "chisq4")) %>%
    ggplot(aes(x = Method, y = mean_value, color = Method)) +
    geom_point() +
    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se)) + 
    facet_grid(. ~ dist) +
    scale_y_continuous("TNR", labels = scales::percent) +
    scale_color_manual(values = col, guide = FALSE) +
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
    ggtitle("TNR in Null Simulations (over 100 replications)") 
gp_nullB

ggsave(file.path(outdir, "FigureSsim-nullA.pdf"), gp_nullA, width = 8, height = 5)
ggsave(file.path(outdir, "FigureSsim-nullB.pdf"), gp_nullB, width = 8, height = 5)

gp_zoom <- plot_grid(pfdr + coord_cartesian(c(0, .2), c(0, .2)) + guides(color = FALSE, linetype = FALSE),
                     ptpr + coord_cartesian(c(.6, 1), c(.6, 1)), rel_widths = c(1, 1.4), labels = LETTERS[1:2])
ggsave(file.path(outdir, "FigureSsim-null.pdf"), gp, width = 8, height = 5)

```

# Informative and Non-Informative Covariates

We subset on simulation results that look at informative and non-informative covariates across
different noise distributions.

```{r}
res_covset <- dplyr::filter(res, grepl("informative", setting))
res_covset <- dplyr::mutate(res_covset, dist = factor(dist,
                                                      levels = c("gaussian", "t11", "t5", "chisq4"),
                                                      labels = c(expression("Noise: " * N(0, 1)),
                                                                 expression("Noise: " * t[11]),
                                                                 expression("Noise: " * t[5]),
                                                                 expression("Noise: " * {chi^2}[4]))))
res_covset <- dplyr::mutate(res_covset, setting = factor(setting,
                                                         levels = c("noninformative", "informative-step",
                                                                    "informative-cubic", "informative-sine")))
res_covset <- tidyr::complete(res_covset, setting, Method, dist, key)
```

We use scatterplots to look at the FDR and TPR values.

```{r, fig.width = 11, fig.height = 4.5}
## looking at FDR vs TPR
col <- as.character(res_covset$col)
names(col) <- as.character(res_covset$Method)

## looking at noninformative vs. informative (cubic)
dplyr::select(res_covset, Method, setting, dist, key, mean_value, se) %>%
    dplyr::filter(setting %in% c("noninformative", "informative-cubic"),
                  key %in% c("TPR"),
                  ! Method %in% c("ashq", "bonf")) %>%
    dplyr::mutate(plus1se = mean_value + se, minus1se = mean_value - se) %>%
    dplyr::select(-se) %>%
    tidyr::gather(metric, value, mean_value, plus1se, minus1se) %>%
    tidyr::unite(metric, c("setting", "metric")) %>%
    tidyr::spread(metric, value) %>%
    ggplot(aes(x = noninformative_mean_value, y = `informative-cubic_mean_value`, color = Method)) +
    geom_errorbar(aes(ymin = `informative-cubic_minus1se`, ymax = `informative-cubic_plus1se`), width = 0) +
    geom_errorbarh(aes(xmin = noninformative_minus1se, xmax = noninformative_plus1se), height = 0) + 
    ##geom_point(alpha = 1) +
    scale_color_manual(values = col) +
    facet_wrap( ~ dist, nrow = 1, scales = "free", labeller = labeller(dist = label_parsed)) +
    scale_x_continuous(labels = scales::percent) + 
    scale_y_continuous(labels = scales::percent) + 
    geom_abline(color = "blue", lty = 2, alpha = 1/3) +
    ggtitle("Mean TPR for Informative and Noninformative Simulations") 
```

```{r}
## looking at noninformative vs. informative (cubic)
pfdr <- dplyr::select(res_covset, Method, setting, dist, key, mean_value) %>%
    dplyr::filter(setting %in% c("noninformative", "informative-cubic"),
                  key %in% c("FDR")) %>%
    tidyr::spread(setting, mean_value) %>%
    ggplot(aes(x = noninformative, y = `informative-cubic`, color = Method)) +
    geom_abline(color = "blue", lty = 2, alpha = 1/3) +
    geom_vline(xintercept = 0.10, color = "red", lty = 2, alpha = 1/3) +
    geom_hline(yintercept = 0.10, color = "red", lty = 2, alpha = 1/3) +
    geom_point(alpha = 1) +
    scale_color_manual(values = col) +
    facet_wrap( ~ dist, nrow = 2, scales = "free", labeller = labeller(dist = label_parsed)) +
    scale_x_continuous("FDR w/ Noninformative Covariate", labels = scales::percent, limits = c(0, 1)) + 
    scale_y_continuous("FDR w/ Informative Covariate (cubic)", labels = scales::percent, limits = c(0, 1)) + 
    ggtitle("FDR (Informative and Noninformative Sims)") 

## looking at noninformative vs. informative (cubic)
ptpr <- dplyr::select(res_covset, Method, setting, dist, key, mean_value) %>%
    dplyr::filter(setting %in% c("noninformative", "informative-cubic"),
                  key %in% c("TPR")) %>%
    tidyr::spread(setting, mean_value) %>%
    ggplot(aes(x = noninformative, y = `informative-cubic`, color = Method)) +
    geom_abline(color = "blue", lty = 2, alpha = 1/3) +
    geom_vline(xintercept = 0.10, color = "red", lty = 2, alpha = 1/3) +
    geom_hline(yintercept = 0.10, color = "red", lty = 2, alpha = 1/3) +
    geom_point(alpha = 1) +
    scale_color_manual(values = col) +
    facet_wrap( ~ dist, nrow = 2, scales = "free", labeller = labeller(dist = label_parsed)) +
    scale_x_continuous("TPR w/ Noninformative Covariate", labels = scales::percent, limits = c(0, 1)) + 
    scale_y_continuous("TPR w/ Informative Covariate (cubic)", labels = scales::percent, limits = c(0, 1)) + 
    ggtitle("TPR (Informative and Noninformative Sims)") 
```

We can plot both FDR and TPR together.

```{r, fig.width = 15, fig.height = 7.5}
gp <- plot_grid(pfdr + guides(color = FALSE, linetype = FALSE), ptpr,
                rel_widths = c(1, 1.4), labels = LETTERS[1:2])
gp

ggsave(file.path(outdir, "FigureSsim-noise.pdf"), gp, width = 11, height = 4.5)

gp_zoom <- plot_grid(pfdr + coord_cartesian(c(0, .2), c(0, .2)) + guides(color = FALSE, linetype = FALSE),
                     ptpr + coord_cartesian(c(.6, 1), c(.6, 1)), rel_widths = c(1, 1.4), labels = LETTERS[1:2])
gp_zoom

ggsave(file.path(outdir, "FigureSsim-noise-zoom.pdf"), gp_zoom, width = 11, height = 4.5)
```

# Unimodal Effect Sizes

Next, we subset on simulation results that look at varying unimodal simulation settings
that were used in the simulations of the ASH paper.

```{r}
res_unimodal <- dplyr::filter(res, grepl("varyinges", setting), dist != "bimodal")
res_unimodal <- dplyr::rename(res_unimodal, signal = dist)
res_unimodal <- tidyr::separate(res_unimodal, setting, c("setting", "dist"), by = "-", fill = "right")
res_unimodal$dist[is.na(res_unimodal$dist)] <- "gaussian"
res_unimodal <- tidyr::complete(res_unimodal, setting, Method, dist, signal, key)
res_unimodal <- dplyr::mutate(res_unimodal, dist = factor(dist,
                                                          levels = c("gaussian", "t", "chisq"),
                                                          labels = c(expression("Noise: " * N(0, 1)),
                                                                 expression("Noise: " * t[11]),
                                                                 expression("Noise: " * {chi^2}[4]))))
```

We take a look at the TPR and FDR as function of the informativeness.

```{r, fig.width = 12, fig.height = 7.5}
## looking at FDR vs TPR
col <- as.character(res_unimodal$col)
names(col) <- as.character(res_unimodal$Method)
lty <- as.character(res_unimodal$lty)
names(lty) <- as.character(res_unimodal$Method)

## We don't include the error bars because the SEs are so small
uniplot <- dplyr::select(res_unimodal, Method, dist, signal, key, mean_value, se, col, lty) %>%
    dplyr::filter(key %in% c("FDR", "TPR")) %>%
    dplyr::mutate(plus1se = mean_value + se, minus1se = mean_value - se) %>%
    dplyr::select(-se) %>%
    tidyr::gather(metric, value, mean_value, plus1se, minus1se) %>%
    tidyr::unite(metric, c("key", "metric")) %>%
    tidyr::spread(metric, value) %>%
    ggplot(aes(x = TPR_mean_value, y = FDR_mean_value, color = Method)) +
    ##geom_errorbar(aes(ymin = FDR_minus1se, ymax = FDR_plus1se), width = 0) +
    ##geom_errorbarh(aes(xmin = TPR_minus1se, xmax = TPR_plus1se), height = 0) + 
    geom_point(alpha = 1) +
    scale_color_manual(values = col) +
    geom_hline(yintercept = 0.10, color = "red", lty = 2, alpha = 1/3) + 
    scale_x_continuous("TPR", labels = scales::percent, breaks = seq(0, 1, .2)) + 
    scale_y_continuous("FDR", labels = scales::percent, breaks = seq(0, 1, .2)) + 
    facet_grid(signal ~ dist, labeller = labeller(dist = label_parsed)) + 
    ggtitle("Average FDR and TPR Across Unimodal Settings")
uniplot

ggsave(file.path(outdir, "FigureSsim-unimodal.pdf"), uniplot, width = 12, height = 7.5)
```

# Varying Null Proportions

Next, we subset on simulation results that look at varying the null proportions.

```{r}
res_pi0set <- dplyr::filter(res, setting == "varyingpi0-lowsignal")
res_pi0set <- dplyr::rename(res_pi0set, pi0 = dist)
res_pi0set <- dplyr::mutate(res_pi0set, pi0 = as.numeric(gsub("nullprop", "", pi0)))
res_pi0set <- tidyr::complete(res_pi0set, setting, Method, pi0, key)
```

We notice that under certain settings, the mean and SE FDR values are `NA`s for some methods under some
settings. We take a closer look at these cases.

```{r}
dplyr::select(res_pi0set, setting, Method, pi0, mean_value, key) %>%
    dplyr::filter(key %in% c("rejections", "FDR")) %>%
        tidyr::spread(key, mean_value) %>%
        dplyr::filter(is.na(FDR))
```

All cases correspond to methods and settings where the method rejected 0 tests over all replications.
We deal with these cases by setting these values to zero for plotting purposes.

```{r}
res_pi0set$se[is.na(res_pi0set$mean_value) & res_pi0set$key == "FDR"] <- 0
res_pi0set$mean_value[is.na(res_pi0set$mean_value) & res_pi0set$key == "FDR"] <- 0
```

We take a look at the TPR and FDR as function of the proportion of null hypotheses.

```{r}
## looking at FDR vs TPR
col <- as.character(res_pi0set$col)
names(col) <- as.character(res_pi0set$Method)
lty <- as.character(res_pi0set$lty)
names(lty) <- as.character(res_pi0set$Method)

## including SE to show variability in FDRR
pi0_fdr <- dplyr::select(res_pi0set, Method, setting, pi0, key, mean_value, se, col, lty) %>%
    dplyr::filter(key == "FDR", !is.na(mean_value)) %>%
    ggplot(aes(x = pi0 / 100, y = mean_value, color = Method, lty = Method)) +
    geom_line(alpha = 0.85, position = position_dodge(width = .01)) +
    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width=0.025, alpha=0.5) +
    scale_color_manual(values = col) +
    scale_linetype_manual(values = lty) + 
    geom_hline(yintercept = 0.10, color = "red", lty = 2, alpha = 1/3) + 
    scale_x_continuous("Null Proportion", labels = scales::percent, breaks = seq(0, 1, .1)) + 
    scale_y_continuous("FDR", labels = scales::percent) + 
    ggtitle("Mean FDR Over 100 Replications") +
    theme_classic() +
    theme(axis.title = element_text(face="bold", size = 10),
          plot.title = element_text(face="bold", size = 14))

## including SE to show variability in FDRR
pi0_tpr <- dplyr::select(res_pi0set, Method, setting, pi0, key, mean_value, se, col, lty) %>%
    dplyr::filter(key == "TPR", !is.na(mean_value)) %>%
    ggplot(aes(x = pi0 / 100, y = mean_value, color = Method, lty = Method)) +
    geom_line(alpha = 0.85, position = position_dodge(width = .01)) +
    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width=0.025, alpha=0.5) +
    scale_color_manual(values = col) +
    scale_linetype_manual(values = lty) + 
    scale_x_continuous("Null Proportion", labels = scales::percent, breaks = seq(0, 1, .1)) + 
    scale_y_continuous("TPR", labels = scales::percent) + 
    ##facet_grid(setting ~ .) + 
    ggtitle("Mean TPR Over 100 Replications") +
    theme_classic() +
    theme(axis.title = element_text(face="bold", size = 10),
          plot.title = element_text(face="bold", size = 14))
```

Again, we can plot both FDR and TPR together.

```{r, fig.width = 15, fig.height = 7.5}
gp <- plot_grid(pi0_fdr + guides(color = FALSE, linetype = FALSE),
                pi0_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2])
gp <- plot_grid(gp, get_legend(pi0_tpr), rel_widths = c(1, .125))
gp

ggsave(file.path(outdir, "FigureSsim-nullprop.pdf"), gp, width = 10, height = 5)
```

# Varying Informativeness

Next, we subset on simulation results that look at varying informativeness of the
informative covariate.

```{r}
res_infoset <- dplyr::filter(res, setting == "varyinginfo-lowsignal")
res_infoset <- dplyr::rename(res_infoset, level = dist)
res_infoset <- dplyr::mutate(res_infoset, level = as.numeric(gsub("level", "", level)))
res_infoset <- tidyr::complete(res_infoset, setting, Method, level, key)
```

We take a look at the TPR and FDR as function of the informativeness.

```{r, fig.width = 8, fig.height = 10}
## looking at FDR vs TPR
col <- as.character(res_infoset$col)
names(col) <- as.character(res_infoset$Method)
lty <- as.character(res_infoset$lty)
names(lty) <- as.character(res_infoset$Method)

## not including SE because it's too small
info_fdr <- dplyr::select(res_infoset, Method, setting, level, key, mean_value, se, q10, q90, col, lty) %>%
    dplyr::filter(key == "FDR") %>%
    ggplot(aes(x = level / 100, y = mean_value, color = Method, lty = Method)) +
    geom_line(alpha = 0.85, position = position_dodge(width = .01)) +
    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width=0.025, alpha=0.5) +
    scale_color_manual(values = col) +
    scale_linetype_manual(values = lty) + 
    geom_hline(yintercept = 0.10, color = "red", lty = 2, alpha = 1/3) + 
    scale_x_continuous("Informativeness", labels = scales::percent, breaks = seq(0, 1, .1)) + 
    scale_y_continuous("FDR", labels = scales::percent) + 
    ggtitle("Mean FDR Over 100 Replications") +
    theme_classic() +
    theme(axis.title = element_text(face="bold", size = 10),
          plot.title = element_text(face="bold", size = 14))

info_tpr <- dplyr::select(res_infoset, Method, setting, level, key, mean_value, se, col, lty) %>%
    dplyr::filter(key == "TPR") %>%
    ggplot(aes(x = level / 100, y = mean_value, color = Method, lty = Method)) +
    geom_line(alpha = 0.85, position = position_dodge(width = .01)) +
    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width=0.025, alpha=0.5) +
    scale_color_manual(values = col) +
    scale_linetype_manual(values = lty) + 
    scale_x_continuous("Informativeness", labels = scales::percent, breaks = seq(0, 1, .1)) + 
    scale_y_continuous("TPR", labels = scales::percent) + 
    expand_limits(y = 0) +
    ggtitle("Mean TPR Over 100 Replications") +
    theme_classic() +
    theme(axis.title = element_text(face="bold", size = 10),
          plot.title = element_text(face="bold", size = 14))
```

Again, we can plot both FDR and TPR together.

```{r, fig.width = 15, fig.height = 7.5}
gp <- plot_grid(info_fdr + guides(color = FALSE, linetype = FALSE),
                info_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2])
gp <- plot_grid(gp, get_legend(info_tpr), rel_widths = c(1, .125))
gp

ggsave(file.path(outdir, "FigureSsim-informativeness.pdf"), gp, width = 10, height = 5)
```

We also create a single plot with simulation results for both varying pi0 and varying
informativeness settings.

```{r}
gp <- plot_grid(pi0_fdr + guides(color = FALSE, linetype = FALSE),
                pi0_tpr + guides(color = FALSE, linetype = FALSE),
                info_fdr + guides(color = FALSE, linetype = FALSE),
                info_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:4])
gp <- plot_grid(gp, get_legend(info_tpr), rel_widths = c(1, .2))
gp

ggsave(file.path(outdir, "FigureAlt-Sims.pdf"), gp, width = 10, height = 8)
```
